DP Problems of Interest

### 2012-10-02 docs/roadmap/problems.tex
More Zucker.

\item Backtracking: same as in \nameref{sswat}
\item Visualisation:

@@

% ----------------------------------------------
\newpage
\subsection{Zuker folding}\ol
\item Problem: folding a RNA string $S$ over itself $\left\lfloor |S| / 2 \right\rfloor = n$.
\item Matrices: $M_{n\times n}, V_{n\times n}, B_{n \times n}, {\color{red} F???}$
\item Alphabets:\ul
	\item Input: $\Sigma(S)=\{A,C,G,U\}$.
	\item Cost matrix: $\Sigma(W)=\Sigma(V)=\Sigma(F)=\{0..n\}$
	\item Backtrack matrix: $\Sigma(B)=\{stop, {\color{red} S,W, XXX here} \}$
	\ule
{\color{red}
\item Initialization:\ul
	\item Cost matrices:\ul
		\item XXX: $W$,$V$
		\item $F_{(0)}=0$
	\ule
	\item Backtrack matrix: $B$
	\ule

\item Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l|l}
	W_{(i+1,j)}+b & S\\
	W_{(i,j-1)}+b & W\\
	V_{(i,j)}+\delta(S_i,S_j) & ?V? \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)} & ?k?
\end{array}\right.\\
\\
V_{(i,j)}&=&\min\left\{\begin{array}{ll}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
	V_{(i+1,j-1)}+es(i,j) \\
	VBI(i,j) \\
	\min_{i<k<j-1}\{W_{(i+,k)}+W_{(k+1,j-1)}\} +c
\end{array}\right.\\
\\
 F_{(j)}&=&\min\left\{\begin{array}{ll} F_{(j-1)} \\ \min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
%
\end{array}\right. \text{(Free Energy)}\\\\
VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c
\end{array}\]
\item Backtracking: XXX
\item Visualisation: XXX
\item Optimizations: XXX
}
\ole

@@

\newpage
\subsection{LMS compiler stack}
\textbf{User-language}: Windowing, input sizes, alphabet, recurrence functions, devices available, what to keep in memory (cost, backtrack or both). \\
$\Downarrow$ Conversion (using an existing technique)

>>>>

\item Backtracking: similar to \nameref{sswat} except that you can jump of $k$ cells.
\item Visualisation:

@@

% ----------------------------------------------
\newpage
\subsection{Zuker folding}\ol
\item Problem: folding a RNA string $S$ over itself $\left\lfloor |S| / 2 \right\rfloor = n$.
\item Matrices: $V_{n\times n}, W_{n\times n}, F_n$ (Free Energy),  $BV_{n \times n}, BW_{n \times n}, BF_n$
\item Alphabets:\ul
	\item Input: $\Sigma(S)=\{A,C,G,U\}$.
	\item Cost matrices:\ul
		\item $\Sigma(W)=\Sigma(V)=\{0..z\}$ with $z \le n*b+c$
		\item $\Sigma(F)=\{0..y\}$ with $y\le \min(F_0, z\cdot n)$
		\ule
	\item Backtrack matrices: \ul
		\item $\Sigma(BW)=\{stop, S,W,V,k\}$
		\item $\Sigma(BV)=\{stop, EH, ES, SW, i' \times j', k\}$
		\item $\Sigma(BF)=\{stop, L, k\}$
		\ule
	\ule
\item Initialization:\ul
	\item Cost matrices: $W_{(i,i)}=V_{(i,i)}=0, F_{(0)}=$ energy of the unfolded RNA.
	\item Backtrack matrices: $BW_{(i,i)}=BV_{(i,i)}=BF_{(0)}=stop$.
	\ule
\item Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l|l}
	W_{(i+1,j)}+b & S\\
	W_{(i,j-1)}+b & W\\
	V_{(i,j)}+\delta(S_i,S_j) & V \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)} &k
\end{array}\right\} = BW_{(i,j)}\\
\\
V_{(i,j)}&=&\min\left\{\begin{array}{l|l}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} & EH\\
	V_{(i+1,j-1)}+es(i,j) & ES \\
	VBI(i,j) & i',j' \\
	\min_{i<k<j-1}\{W_{(i+1,k)}+W_{(k+1,j-1)}\} +c & k
\end{array}\right\} = BV_{(i,j)}\\
\\
F_{(j)}&=&\min\left\{\begin{array}{l|l}
	F_{(j-1)} & L \\ 
	\min_{1\le i< j} (V_{(i,j)} + F_{(j-1)}) & k
\end{array}\right\} = BF_{(j)}\\
\\
VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c
\end{array}\]

In practice, we don't go backward for larger values than 30, so we can replace $\min_{i<k<j}$ by $\min_{\max(i,j-30)<k<j}$ in the expressions of $W$,$V$, $VBI$ and $F$.

{\color{red}
\item Backtracking: XXX
\item Visualisation: XXX
\item Optimizations: XXX
}
\ole

@@

\newpage
\subsection{LMS compiler stack}
\textbf{User-language}: define additional parameters for the recurrence\ul
\item Windowing (to convert non-serial into serial problems)
\item Input sizes, and alphabets (backtrack, input, cost)
\item Backtrack (implicitly by backtrack alphabet size) and cost matrices bit-sizes (cost maximum may be inferred using <<Yield size/grammar analysis>>)
\item Recurrence functions, devices available
\item What to keep in memory (cost, backtrack or both).
\ule
$\Downarrow$ Conversion (using an existing technique)
